Transition path sampling based calculations of free energies for enzymatic reactions: the case of human methionine adenosyl transferase and plasmodium vivax adenosine deaminase

### 2022-06-07 tex/resubmission1/response/response.tex

\pdfoptionpdfminorversion=4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{chemformula} % Formula subscripts using \ch{}
\usepackage[T1]{fontenc} % Use modern font encodings
\usepackage{graphicx}
\usepackage{caption}
\usepackage{amsmath}
\usepackage{mathptmx}
\usepackage[scaled=0.92]{helvet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\usepackage{microtype,xparse,tcolorbox}
\usepackage{helvet}
\newcommand{\newtext}[1]{{\color{blue} #1}}
\newcommand{\vamsee}[1]{{\color{applegreen} #1}}
\newcommand{\filipp}[1]{{\color{red} #1}}

\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~}
      \setlength\parindent{2em}
      \noindent
      \ttfamily #1
    \end{reviewer-comment~}
  }
  \par\noindent\ignorespaces
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}
\ExplSyntaxOff
%\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
%\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
%\newcommand\PaperTitle[1]{``\textit{#1}''}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Sree Ganesh Balasubramani}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
\author{Steven D. Schwartz}
\affiliation{Department of Chemistry and Biochemistry, University of Arizona, Tucson, Arizona 85721, United States}
%\author{I. Ken Groupleader}
%\altaffiliation{A shared footnote}
\email{sschwartz@email.arizona.edu}
%\phone{+123 (0)123 4445556}
%\fax{+123 (0)123 4445557}
%\affiliation[Unknown University]
%{Department of Chemistry, Unknown University, Unknown Town}
%\alsoaffiliation[Second University]
%{Department of Chemistry, Second University, Nearby Town}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {Transition path sampling based calculations of free energies for enzymatic
  reactions: the case of human methionine adenosyl transferase and plasmodium 
  vivax adenosine deaminase}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{TPS,WHAM}
\keywords{American Chemical Society, \LaTeX}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%\date{\today}

\maketitle

%Additions to the manuscript are indicated in  \textit{\newtext{blue text}}.

\Reviewer{1}

We would like to thank the Reviewer for his/her time and for evaluating
the manuscript.

%\begin{response}
%{The current work shows the application of the BOLAS method to two enzyme-catalyzed reactions. 
%The BOLAS method is a transition path sampling algorithm that can be extended to calculate free 
%energies along a pre-defined reaction coordinate. There are already several methods to calculate 
%free energy profiles and many have been applied to enzymatic reactions, but none is a black-box, 
%and therefore the implementation and application of this method to an enzymatic reaction is an 
%important advance. The text is well written and clear.}
%reply
%TPS based free energy calculations have been attempted previously for the calculations
%involving model problems involving analytical potential energy surfaces or for 
%configurational or structural changes of an enzyme. A chemical reaction catalyzed by 
%an enzyme can involve a complicated reaction coordinate and TPS based method using the 
%shooting algorithm provides a bias free technique for calculating the free energy
%of the reaction.

%\end{response}

\begin{response}
{First one needs to know how the specific parameters of the method affect 
the results. Or at least, how or why they were chosen. If not for all parameters, 
at least for the most relevant ones (but this needs also to be specified). 
For example:
* why use 150 reactive trajectories?}
%reply

%The application of TPS for stuyding the kinetics and transition states of 
%enzyme catalyzed reactions has been repeorted several times by our group as well
%as by others we decided to not list out every detail. But we have now included 
%some details about the choice the number of trajectories and so on.
The number of reactive trajectories in a TPS ensemble is decided based 
on how quickly the shooting algorithm results in decorrelation of the reactive 
trajectories from the initial biased trajectory. If the decorrelation is fast 
then fewer trajectoreis are sufficient for further committor analysis. 
To be sure that we do not use trajecotries that are correlated to the initial 
biased trajectory we ignore the first 100 trajectories for committor analysis. 
This choice for the number of trajectories is also based on computational cost, 
while obtaining the additional trajectories may not provide any more information 
we use up computational resources in the process.

\end{response}

\begin{response}
  {why should each reaction last 0.5 ps? How is it known if that time is too short 
  or too long?}
We used longer trajectories (1 ps) and checked to make sure that the reaction occurs within
the time that we used (0.5 ps) before collecting the TPS ensemble.

\end{response}

{\begin{response}
{* why 20 windows?}
%reply
This choice was made based on the order parameter window that we are scanning.
20 windows results in a window size of 0.39 angstroms and we use a overlap of
0.09 angstroms on either side of the window to make sure that we have configurations
that are common to neighboring windows which is necessary to obtain a continuous free energy
profile. Smaller window sizes result in inefficient sampling because the TPS based method
does not include bias forces to keep the trajectories within the windows and we end up
sampling outside the windows especially in regions where the free energy has a steeper slope.

% We are somewhat at loss how to respond to this criticism,
% which seems to contradict Reviewer B's statements,
% ``I have to stress that the paper discusses complex material and still
% it is in general clearly written. It can be accessible for readers who
% are less familiar with RPA or many-body theory.''

\end{response}

\begin{response}
{* When calculating the commitor, you cannot say "repeating this procedure many times". 
How many are needed to obtain a given precision or to obtain the given figure?
\newline
* As before, when calculating the separatrix, you can not say "repeating this 
process for several trajectories".}
%reply
Since this is the discussion of the committor analysis and committor distribution 
analysis in general and not for the specific case for the enzymes we had not
provided the number of trajectories that were used in these calculations.

\end{response}

\begin{response}{Some of these questions arise because nowhere in 
the manuscript are confidence intervals given. Neither in the energy 
profile nor in the transition state structure distances (Fig. 5). 
Therefore the effect of these parameters on the precision of the 
results is not known.
}
The transition state structure depicted in Fig. 5 is a representative
from the ensemble of transition states that were obtained from committor
analysis. We could provide the mean and standard deviation of the bond distances
but this does not provide the any more

\end{response}

\begin{response}{Some of these questions also arise because the 
code to implement the method is not made public. Doing so would 
not only clarify some questions but also prevent others from 
reinventing the wheel, that is, coding the implementation again 
to check the results of this paper or to apply the method to their specific enzymes.}

% REPLY
We agree that the availability of code is necessary for reproducibility of the results.
TPS based free energies calculations are akin to vanilla TPS calculations with 
only a few modifications to the acceptance criteria as discussed in the manuscript as well
as in the original BOLAS paper. We followed the algorithm stated in the manuscript
and implemented it as python scripts but it was hard coded to model the MAT2A and 
$\textit{pv}$ADA catalyzed reactions. We would release it in a public repository once it is 
tailored for general purpose usage.  
\end{response}

\begin{response}{The second piece of information needed is how the method 
performs compared to others. Both in terms of computational cost and in 
terms of accuracy. Comparison with experiment is interesting but not very 
enlightening as the computational results depend on the method to calculate 
the free energy profile (tested here) but also on the chosen reaction coordinate, 
the QM region and the QM method, among others. A comparison with Umbrella Sampling, 
which also uses a reaction coordinate and also uses windows would be ideal. Again, 
to make fair comparisons we would need error bars to estimate the convergence of 
the results and reports of computational costs.}

% REPLY

\end{response}

\begin{response}{Beyond these general issues raised so far, there are some other points 
that deserve some attention.
\newline
1. How is the initial reactive trajectory generated? Although obtaining an initial reactive 
trajectory is not the crux of the algorithm it is a non-trivial step that should be more 
detailed. In particular, for the pvADA enzyme, which involves a two-step reaction, how 
do the way these 2 steps take place in the initial trajectory condition of the TPS ensemble?}

% REPLY
For the \textit{pv}ADA enzyme the two steps in the reaction are the nucleophilic attack 
of the OH$^{-}$ on the C6 atom and the transfer of the proton from the OH$^{-}$ 
to the N6 atom with the formation of NH$_3$. We are including the time series of the 
N6-H1 distances in the supporting information which would show the sequence of events 
along a typical TPS trajectory for this reaction.

\end{response}

\begin{response}{2. Related to the previous question, considering that the barriers 
surrounding the intermediate are higher than kT, shouldn't the trajectories remain 
trapped in the intermediate? In other words, isn't the duration of the whole trajectory 
determining the mechanism by enforcing a ballistic trajectory? Is the d(NC)-d(OC) still 
a good reaction coordinate to describe the second step? If not, shouldn't that underestimate 
the barrier, something that should be compensated by a (not calculated) transmission coefficient?}

% REPLY

\end{response}

\begin{response}
{3. I find the free energy profiles in fig 7 and 11 worrying. Take the 4th window in 
Fig. 7: it is clearly showing a smaller slope than the interpolation and this is because 
there are jumps between the windows. The first point of this window and the next one are 
"disconnected" from the other ones. Why is this so? Repeating the calculations with smaller 
windows would show the effect of window size on the result (that would address one of my 
questions regarding the choice of parameters).}
% REPLY
\end{response}

\begin{response}
{4. The author should check they cite the appropriate works. I am afraid 
ref 20 is not an application of TPS to enzyme catalysis. To my knowledge, The first 
applications of TPS to enzyme catalysis are  $J. Am. Chem. Soc.2005, 127, 13822-13831$ 
and $J. Phys. Chem. B 2007, 111, 20, 5708-5718$.}
% REPLY

We thank the Reviewer for pointing out this mistake. We have made the appropriate changes to that
citation in the manuscript. 
\end{response}

\begin{response}{Overall this is a relevant work, but the authors need to go 
deeper into the scientific details to be useful to other members of the 
community, because what the manuscript should be describing is not that the 
method works --we already knew that-- but how to make it work in enzymatic reactions.}
% REPLY
\end{response}

\section*{Other Changes}

\end{document}

### 2022-06-07 tex/resubmission1/supp-info/si.tex
added the new files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=article,layout=traditional]{achemso}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}

%Some journals require a graphical entry for the Table of Contents.
%This should be laid out ``print ready'' so that the sizing of the
%text is correct.

%Inside the \texttt{tocentry} environment, the font used is Helvetica
%8\,pt, as required by \emph{Journal of the American Chemical
%Society}.

%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
%permitted for  \emph{Journal of the American Chemical Society}
%graphical table of content entries. The box will not resize if the
%content is too big: instead it will overflow the edge of the box.

%This box and the associated title will always be printed on a
%separate page at the end of the document.

%\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  Transition path sampling (TPS) is widely used for the
  calculations of reaction rates, transition state structures and reaction 
  coordinates of condensed phase systems. Here we discuss a scheme 
  for the calculation of free energies using the ensemble of TPS reactive 
  trajectories in combination with a window based sampling technique 
  for enzyme catalyzed reactions. We calculate the free energy profiles of 
  the reactions catalyzed by the human methionine S-adenosyltransferase (MAT2A) 
  enzyme and the plasmodium vivax adenosine deaminase ($pv$ADA) enzyme to assess 
  the accuracy of this method. MAT2A catalyzes the formation of S-adenosine-L-methionine
  following a S$_{\text{N}}2$ mechanism and using our method we estimate
  the free energy barrier for this reaction to 
  be $16\;\text{Kcal}\;\text{mol}^{-1}$ which
  is in excellent agreement with the experimentally measured activation 
  energy of $17.27\;\text{Kcal}\;\text{mol}^{-1}$. Furthermore, for the 
  $pv$ADA enzyme catalyzed reaction we estimate a free energy barrier of 
  $23\;\text{Kcal}\;\text{mol}^{-1}$ and the calculated free energy profile is
  similar to that predicted from experimental observations.
  Calculating free energies employing our simple method within TPS 
  provides significant advantages over such methods as umbrella sampling 
  since it is free from any applied external bias, accurate compared to 
  experimental measurements and has a reasonable computational cost.    
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimates of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecules in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades. These include 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,
Dellago98JChemPhys108p1964,Bolhuis21AdvTheory4p2000237}

TPS is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS is based on Monte Carlo importance sampling in the 
space of trajectories. It uses the shooting and shifting algorithm 
to collect an ensemble of reactive pathways that are real dynamical 
trajectories generated without the application of any bias forces. \cite{dellago02AdvChemPhys123} 
Therefore the TPS ensemble can be used to obtain 
kinetic information such as the rate of the reaction as well as to elucidate 
the transition state, reaction coordinate and the reaction mechanism. 
TPS has been successfully applied to study problems such as enzyme catalysis, 
\cite{Antoniou01JPhysChemB105p5553,Hay12NatChem4p161,Schwartz09NatChemBiol5p551} 
and protein folding. \cite{Bolhuis03ProcNatlAcadSci100p12129,Juraszek12ChemPhys396p30}

Calculating equilibrium properties such as the free energy of an enzymatic 
catalysis reaction within TPS is not straightforward. 
The Monte Carlo algorithm used in TPS has an acceptance criterion that 
is only satisfied if the trajectory begins in the reactant basin and ends in 
the product basin or vice versa, hence trajectories that are localized in the 
reactant or product basins, without visiting both are rejected. 
If one defines an order parameter $\zeta$, which separates reactant from 
product basins, then all trajectories, reactive and non-reactive, 
contribute to the probability density distribution function 
($P(\zeta)$) which is the probability of a system at equilibrium to have 
a particular value of the order parameter. \cite{Dellago09AdvCompSimAppp167}

Frenkel has shown that inclusion of rejected states in a Markov state Monte 
Carlo algorithm can help speed up the convergence of the calculations of 
the order parameter distribution of a two-dimensional
Ising model. \cite{Frenkel04ProcNatAcadSci101p17571}
Inclusion of rejected trajectories within TPS was first discussed by Radhakrishnan 
et al. \cite{Radhakrishnan04JChemPhys121p2436} who developed `BOLAS' which combines 
the shooting and shifting algorithms of TPS with a window based sampling technique 
in the spirit of umbrella sampling to calculate equilibrium free energy 
differences. Peters et al. used a variant of the BOLAS scheme called the 
equilibrium path sampling (EPS) where they use aimless shooting 
combined with window based sampling for free energy calculations. 
\cite{Peters08JAmChemSoc130p17342,Beckham10epsbook} 
More recently Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}
have developed a re-weighted path ensemble scheme within the standard TPS method 
called virtual interface exchange TPS (VIE-TPS) to calculate the free energy 
landscape. VIE-TPS is shown to produce reasonably accurate free energy profiles for 
chemical reactions that are diffusive in nature whereas it is less accurate otherwise.
Transition interface sampling (TIS) has been used to obtain 
a re-weighted path ensemble which can give access to free energies 
and committor surfaces. \cite{Rogal10JChemPhys17p174109}

Free energy calculations based on the TPS shooting algorithm is 
appealing but rare in the literature. Applications of this 
method to study enzymatic reactions and comparisons of the calculated 
free energies to experimental measurements can shine light on the 
accuracy of this approach compared to the standard methods used for 
free energy calculations.    
In this manuscript we implement and apply the 
algorithm developed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
to calculate the free energies of reactions catalyzed by two systems:
the human MAT2A enzyme and the plasmodium vivax adenosine deaminase 
enzyme. \cite{Luo07JAmChemSoc129p8008,Ho09Biochemistry48p9618}
Human MAT2A catalyzes the formation of S-adenosyl L-methionine (SAM).
SAM is an essential metabolite which is 
distributed to almost all body tissues and fluids, is a universal 
methyl donor and is of fundamental importance to the metabolism of 
compounds such as hormones, neureotransmitters, proteins, 
and nucleic acids. \cite{Friedel89Drugs38p389}
Plasmodium vivax is a parasite that is responsible for the largest number 
of cases of malaria globally. Targeting proteins that are essential to
the metabolism of this bacteria such as the adenosine deaminase enzyme 
is a possible anti-malarial drug development strategy. \cite{Madrid08JBiolChem283p35899} 
These enzymes were chosen not only because of their importance in 
biochemical applications but also because of the availability of experimental 
results with which comparisons can be drawn. This manuscript is organized as 
follows: first we discuss the TPS method and the algorithm that we use 
to calculate the free energies, then we show calculations which apply this 
method on the two enzymatic reactions. Finally we compare our results to experiments
and provide conclusions. 
%------------------------------------------------------------------------------ 
\section{Methods}
%------------------------------------------------------------------------------ 
\subsection{Transition path sampling}
%------------------------------------------------------------------------------ 
Here we briefly describe the equations describing the TPS method for the purpose of
introducing notations as well as for the sake of completeness. 
Consider the molecular dynamics simulation of a molecular system consisting 
of $N$ atoms. Starting from the initial positions and momenta of each atom at time $t_0$,
the time evolution is carried out by finding the position ($\textbf{q}$) and 
momenta ($\textbf{p}$) of each atom at regular intervals of time $t_i = t_0 + i\Delta t$  
($i = 0,1,2,\ldots$). At each of the time slices, the positions and 
momenta can be collectively represented as
the set $z = \{\textbf{q},\textbf{p}\}$. If the MD simulation is 
run for a total 
time of $\mathcal{T}$ the number of time slices is given by 
$L = \mathcal{T}/\Delta t +1$ and for this sequence of 
times the state of the system can be represented as 
\begin{equation}
z(\mathcal{T}) = \{z_0, z_{\Delta t}, z_{2\Delta t},\ldots,z_{\mathcal{T}}\}
\end{equation}
The probability to obtain a particular sequence of states is determined by the initial 
conditions and the type of dynamics used for the time evolution. If the time evolution is 
Markovian, the probability to go from $z_{i\Delta t}$ to $z_{(i+1)\Delta t}$ depends only 
on $z_{i\Delta t}$ and not on conditions before the time $i\Delta t$ the total path probability can 
be expressed as the product of individual probabilities $p(i\rightarrow j)$ as 
\begin{equation}
\mathcal{P}[z(\mathcal{T})] = \rho(z_0)\Pi_{i=0}^{\mathcal{T}/\Delta t-1} p(z_{i\Delta t}\rightarrow z_{(i+1)\Delta t}),
\end{equation}
where $\rho(z_0)$ is the equilibrium distribution of the initial conditions and for a canonical ensemble this can be 
expressed as 
\begin{equation}
\rho(z_0) = \exp(-\beta H(z_0))/Z
\end{equation}
where 
\begin{equation}
Z = \int dz \exp(-\beta H(z)) 
\end{equation}
is the canonical partition function.

Within the transition path sampling, the trajectories of interest start in the reactant region of the 
phase space ($\mathcal{A}$) and end in the product region ($\mathcal{B}$). Such reactive trajectories have 
a restricted probability distribution function given by
\begin{equation}
\mathcal{P}_{\mathcal{AB}}[z_{\mathcal{T}}] = h_{\mathcal{A}}(z_0)\mathcal{P}[z(\mathcal{T})]
h_{\mathcal{B}}(z_{\mathcal{T}})/Z_{\mathcal{AB}}(\mathcal{T})\label{eqn:tpsensem}
\end{equation}
where 
\[
    h_{\mathcal{A}/\mathcal{B}}(z)= 
\begin{cases}
    1, & \text{if } z\in \mathcal{A}/\mathcal{B}\\
    0,              & \text{otherwise}
\end{cases}
\]
and $Z_{\mathcal{AB}}(\mathcal{T})$ is the normalization factor for this 
probability distribution function.
The set of all reactive trajectories characterized by the probability 
distribution function given by Eq. \ref{eqn:tpsensem} is the 
transition path ensemble.

\subsection{Equilibrium distribution of order parameters}
The free energy of a molecular system as a function of the order parameter $\zeta$ 
is defined as 
\begin{equation}
\text{A}(\zeta) = -k_{\text{B}}T\text{ln}(P(\zeta)) + C, \label{eqn:fenergy}
\end{equation}
where $P(\zeta)$ is the probability distribution function of the reaction coordinate
$\zeta$, $C$ is an arbitrary constant, $k_{\text{B}}$ is the Boltzmann constant
and $T$ is the temperature. At equilibrium the probability distribution of the 
reaction coordinate can be obtained from the distribution function for the phase 
space $\rho(\textbf{q})$ as
\begin{equation}
P(\zeta) = \int d\textbf{V} \rho(\textbf{q})\delta\left[\zeta-\tilde{\zeta}(\textbf{q})\right].
\end{equation}
The integration is over the entire phase space. In practice, the calculation of this distribution function
often proceeds using histogram based methods where the reaction coordinate is 
divided into bins and the frequency of the occurrence of the reaction coordinate 
within a particular window during the course of MD simulation is used to calculate 
the probability distribution function. Since some values of the reaction coordinates 
particularly close to the transition state are 
rarely sampled in a conventional MD simulations, enhanced sampling techniques such as the umbrella sampling
are necessary to sample the low probability regions. TPS is designed to sample these regions in phase space 
without the application of any external bias, but the trajectories 
sampled with TPS are only reactive and hence
they are not distributed according to the equilibrium distribution function. 
The restriction that the pathways start from the reactant state and end in the 
product state can be relaxed to obtain an ensemble of trajectories that will 
be distributed according to the equilibrium distribution. 
%------------------------------------------------------------------------------ 
\subsection{Algorithm for free energy calculations}\label{ssec:algorithm}
%------------------------------------------------------------------------------ 
We follow the steps proposed by Radhakrishnan et al. \cite{Radhakrishnan04JChemPhys121p2436} 
to sample TPS trajectories within windows of the order parameter. 
\begin{itemize}
\item An appropriate order parameter is chosen for the reaction of interest 
that can distinguish the reactant and product states sufficiently well and the TPS ensemble is harvested 
which consists only of reactive trajectories.
\item The order parameter is divided into windows $\{\zeta_i\}$ and within each window,
$\zeta_{i}^{min} < \zeta < \zeta_{i}^{max}$ is to be sampled.
\item Choose a reactive trajectory $\tilde{z}(\mathcal{T})$ from the TPS ensemble as 
a guiding trajectory for the window sampling. Pick a time slice from this trajectory 
$\tilde{z}_{j\Delta t}$ such that the order parameter calculated for this time slice
$\tilde{\zeta}$ satisfies $\zeta_{i}^{min} < \tilde{\zeta} < \zeta_{i}^{max}$.
\item Using the shooting and shifting algorithm familiar from TPS simulations, \cite{dellago02AdvChemPhys123}
harvest dynamics trajectories starting from the time slice $\tilde{z}_{j\Delta t}$ of the trajectory 
$\tilde{z}(\mathcal{T})$ and accept the new trajectory if for any time slice the calculated 
order parameter $\hat{\zeta}$ satisfies $\zeta_{i}^{min} < \hat{\zeta} < \zeta_{i}^{max}$.
\item Calculate the probability distribution ($P_i(\zeta)$) within the window $\zeta_{i}^{min} 
< \zeta < \zeta_{i}^{max}$ by constructing histograms obtained from the ensemble of 
accepted trajectories for the window referenced by the index $i$. 
\item Calculate the free energies for each of the windows according to Eq. \ref{eqn:fenergy}
and combine them such that the function $\text{A}(\zeta)$ is continuous 
by adjusting the constants $C$. 
\end{itemize}
%------------------------------------------------------------------------------ 
\section{Computational details}
%------------------------------------------------------------------------------ 
\subsection{System preparation}
%------------------------------------------------------------------------------
We used the crystal structure of the human MAT2A enzyme 
bound to methylthioadenosine and di-imido triphosphate (PNPNP)
(protein data bank ID: 7L1A) reported by Niland 
et al. \cite{Niland21Biochem60p791,Ghosh21JAmChemSoc143p18325} and the crystal 
structure of \textit{pv}ADA in complex with MT-coformycin (protein data bank ID: 3EWC) 
reported by Ho et al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
simulations. For calculations involving the MAT2A enzyme the substrates were modified to 
L-methionine and adenosyl triphosphate (ATP) whereas for calculations with the \textit{pv}ADA
enzyme the substrates were modified to $2^{'}-$deoxyadenosine and hydroxide ion. 
We solvated both the systems using TIP3P water molecules \cite{Jorgensen83JChemPhys79p926}
in a nanodroplet sphere whose volume was fixed at $15$ {\AA}
away from the protein's surface and the total charge was neutralized 
using potassium ions.

MD simulations within classical and hybrid quantum mechanics/molecular mechanics 
(QMMM) approximations were carried out using the CHARMM program 
package. \cite{Brooks83JComputChem4p187} 
For the MM calculations the CHARMM36 force field \cite{Brooks09JComputChem30p1545} 
was used and for the QM calculations the approximate PM3 semiempirical 
method \cite{Repasky02JComputChem23p1601} was used. 
For the hybrid QMMM simulations, the MAT2A system was partitioned into a QM
region consisting of the ATP and MET molecules along with three Mg$^{2+}$ ions 
and the MM region consists of the rest of the molecules. 
For the $pv$ADA system the QM region consisted of 
the adenosine molecule, an hydroxide anion, Zn$^{2+}$ ion and 
a portion of the Glu229 residue which was partitioned within the generalized 
hybrid orbital (GHO) scheme. \cite{Gao98JPhysChemA102p4714}

The energies of both the systems were minimized using 50 steps of the 
steepest descent method, followed by 2000 steps of the
adopted basis Newton-Raphson method where only classical molecular mechanics 
was used for the dynamics. 
The minimized systems were
heated slowly to 300 K for 35 ps beginning with harmonic
constraints on all atoms except on the H atoms and the TIP3P 
water molecules with gradual reduction of the restraint forces. 
15 ps of equilibration was carried out starting with harmonic 
restraint forces followed by 20 ps of constraint free 
equilibration to prepare the systems for TPS simulations. 
During the heating and equilibration steps, bonds that contain 
hydrogen atoms in the MM region were restrained to their equilibrium values 
using the SHAKE procedure. \cite{Ryckaert77JComputPhys23p327} 
%------------------------------------------------------------------------------ 
\subsection{Transition path sampling}
%------------------------------------------------------------------------------
TPS simulations require the definition of unique reactant and product states
to collect reactive trajectories based on the Monte Carlo algorithm. Here we 
discuss the details of the reactions involving the MAT2A and the $pv$ADA enzymes 
to identify and define the order parameters characterizing the reactant and product basins.

The reaction catalyzed by the human MAT2A enzyme is
between adenosyl $5'$-triphosphate (ATP) and L-methionine (MET) molecules 
resulting in the formation of S-adenosyl-L-methionine (SAM) which is 
depicted in Fig. \ref{fig:mat2a-reaction}. This reaction follows a S$_{\text{N}}2$ 
mechanism and was experimentally characterized by 
Firestone et al. \cite{Firestone17JAmChemSoc139p13754}. The key bond 
parameters for this reaction consists of the distance between the
nucleophilic sulfur atom of MET and the $5'-C$ atom of the ATP (denoted as d$_{\mathrm{SC}}$), 
and the distance between oxygen atom of the phosphate group and the $5'-C$ 
atom of ATP (d$_{\mathrm{OC}}$). For the TPS calculations we define the 
reactant state to have d$_{\mathrm{OC}}$ to be less than 1.7 {\AA} and the product 
state to have the d$_{\mathrm{SC}}$ less than 2.0 {\AA}.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/mat2a-reaction.png}
\caption{The reaction catalyzed by the human MAT2A enzyme between ATP and L-methionine resulting in the 
formation of S-adenosyl L-methionine and triphosphate leaving group.}
\label{fig:mat2a-reaction}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jpcbfk,manuscript=suppinfo,layout=traditional]{achemso}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{tocentry}
\section{Initial reactive trajectory}
\begin{figure}
\includegraphics[scale=1]{figures/ada-oh-nh60.pdf}
\caption{The distance between the N6-H1 atoms and O1-H1 atoms as a function of time. The proton transfer from the hydroxide
anion to the NH$_2$ group occurs after the nucleophilic attack of the O1 atom on the C6 atom.}
\end{figure}
